Parameter Space Representation of Pareto Front to Explore Hardware-Software Dependencies

### 2014-10-30 introduction.tex
Final english fixes

In this work we present PS, a multiobjective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front, i.e. how different these configurations are, in terms of objectives, with respect to the ones already found previously. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

@@

The paper is organized as follows. First, \secR{Related-work} places our work in the context of related effort.
Then we give a theoretical formulation of the algorithm. In particular, in \secR{statement_of_the_problem} we provide formal definitions of the concepts and the operations that will be used to describe PS algorithm. In \secR{algorithm} we give the theoretical description of PS. 
Finally in \secR{ee} we show a case
study involving the exploration of the hardware/software parameters of
a Very Long Instruction Word (VLIW) architecture, performing a
qualitative and quantitative comparison of PS against the state-of-art
of multiobjective genetic based approaches.

\section{Related work and Our Contribution}
\secL{Related-work}
Many different design space exploration algorithms have been proposed
in literature.  What makes the exploration of embedded systems so
complex is not only the huge size of the parameter spaces involved,
but also the tight interaction between parameters in relation to their
effect on the objective to be optimised, which take the form of
\emph{parameter dependencies} (see~\defR{dependency}). The 
motivations of an exploration algorithm and how it deals with such
complexity are often rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

A preliminary estimation of hardware/software dependencies in order to
reduce the design space is addressed in~\cite{Catania2008}, where
authors propose to deal with the software parameters separately,
statistically testing their average effect on objectives before the
actual exploration is performed. Also in~\cite{merging} authors
analyze the impact of including compilation parameters into the design
space space exploration. The results of both works confirm the
important role that hardware/software dependencies can play in a
multi-objective exploration and thus justify the choice of explicitly
including the two class of parameters in the design space without
making any distinction between them from the algorithm's perspective.
In fact, one of the purposes of the proposed PS algorithm is avoiding
the need of any a priori knowledge about those dependencies, letting
the algorithm itself discover regions where interaction between
parameters generates new pareto-optimal points (see the PS scoring
system described in the next sections).

@@

The most well documented and widespread class of algorithms adopted in the
design space exploration of embedded systems is probably the one of Evolutionary
Multiobjective Optimization Algorithms~\cite{coello_easmop}. For this
reason this is the class of algorithms that we will use for the
experiments discussed in~\secR{ee}.
Evolutionary approaches have many advantages: they provide good accuracy, they are
problem-independent and do not require any a-priori knowledge of the
system. The strong point of these approaches is that they consist in an exploration that is sufficiently broad
(the mutation avoids being rigidly restricted to limited parts of
parameter space), and, at the same time, not too scattered as it is
guided by the performances of the already evaluated configurations.
Different variants of these algorithms have been proposed in
literature, including mixed approaches as in~\cite{Ascia2011382},
where authors specifically focuse on increasing the performance of a
evolutionaly exploration strategy, proposing a fuzzy based technique
to avoid the actual simulation of a certain number of configurations.
In particular, in this work we will compare the proposed approach against a
well known SPEA2 multi-objective genetic algorithm~\cite{knowles_techrep06}
which is, at best of our knowledge, one of the most appropriate and
tested for the particular scenario we are assuming.  We
claim that the approach presented in this paper has most of the
benefits of the evolutionary approaches, although the rationale is
completely different.

>>>>

In this work we present PS, a multi-objective design exploration
algorithm that introduces the concept of \emph{interesting} and
\emph{uninteresting regions} of the configuration space.  The main
idea can be simplified as follows: at each iteration of the algorithm,
the configuration space is divided in regions. A score is given to
each region based on the \emph{innovation} provided by its
configurations in the Pareto front, i.e. how different these configurations are, in terms of objectives, with respect to the ones already found previously. The regions with high score are considered the most \emph{interesting} and more simulations are distributed there rather than in the rest of the configuration space.
Several geometrical transformations (splitting, merging etc.) are
involved in the Parameter Space (aka configuration space), hence
the name PS for the algorithm.

@@

The paper is organized as follows. First, \secR{Related-work} places our work in the context of related effort.
Then we give a theoretical formulation of the algorithm. In particular, in \secR{statement_of_the_problem} we provide formal definitions of the concepts and the operations that will be used to describe PS algorithm. In \secR{algorithm} we give the theoretical description of PS. 
Finally in \secR{ee} we show a case
study involving the exploration of the hardware/software parameters of
a Very Long Instruction Word (VLIW) architecture, performing a
qualitative and quantitative comparison of PS against the state-of-art
of multi-objective genetic based approaches.

\section{Related work and Our Contribution}
\secL{Related-work}
Many different design space exploration algorithms have been proposed
in literature.  What makes so complex the exploration of embedded
systems is not only the huge size of parameter spaces involved,
but also the tight interaction between parameters in relation to their
effect on the objective to be optimized, which take the form of
\emph{parameter dependencies} (see~\defR{dependency}). The 
motivations of an exploration algorithm and how it deals with such
complexity are often rather heuristic, have some form of arbitrariness and,
to a large extend, intuition lies behind them.

@@

A preliminary estimation of hardware/software dependencies in order to
reduce the design space is addressed in~\cite{Catania2008}, where
authors propose to deal with the software parameters separately,
statistically testing their average effect on objectives before the
actual exploration is performed. Also in~\cite{merging} authors
analyze the impact of including compilation parameters into the design
space exploration. The results of both works confirm the
important role that hardware/software dependencies play in a
multi-objective exploration and thus justify the choice of explicitly
including the two class of parameters in the design space without
making any distinction between them from the algorithm's perspective.
In fact, one of the purposes of the proposed PS algorithm is to avoid
the need of any a priori knowledge about those dependencies, letting
the algorithm itself to discover regions where interaction among
parameters generates new Pareto-optimal points (see the PS scoring
system described in the next sections).

@@

The most well documented and widespread class of algorithms adopted in the
design space exploration of embedded systems is probably the Evolutionary
Multi-objective Optimization Algorithms~\cite{coello_easmop}. For this
reason, this is the class of algorithms that we will use for the
experiments discussed in~\secR{ee}.
Evolutionary approaches have many advantages: they provide good accuracy, they are
problem-independent and do not require any a-priori knowledge of the
system. The strong point of these approaches is that they consist in an exploration that is sufficiently broad
(the mutation avoids being rigidly restricted to limited parts of
parameter space), and, at the same time, not too scattered as it is
guided by the performances of the already evaluated configurations.
Different variants of these algorithms have been proposed in
literature, including mixed approaches as in~\cite{Ascia2011382},
where authors specifically focus on increasing the performance of an
evolutionary exploration strategy, proposing a fuzzy-based technique
to avoid the actual simulation of a certain number of configurations.
In particular, in this work we will compare the proposed approach against a
well known SPEA2 multi-objective genetic algorithm~\cite{knowles_techrep06}
which is, at best of our knowledge, one of the most appropriate and
tested for the particular scenario we are assuming.  We
claim that the approach presented in this paper has most of the
benefits of the evolutionary approaches, although the rationale is
completely different.

### 2014-10-30 paper.tex
AA: altre modifiche
AA: ultimati i TODO AND sul testo
added platform description
AA
Final english fixes

@@

%\maketitle
\begin{abstract}
%Embedded Systems Design requires conflicting objectives to be
%optimized with an appropriate choice of hardware/software parameters.
%In this work we present PS, a new multiobjective exploration algorithm
%exploiting the parameter space representation of the pareto front.
%After a detailed formal description of the algorithm and the
%underlying concepts, we show a case study involving the
%hardware/software exploration of a VLIW architecture. Qualitative and
%quantitative comparisons of the PS approach against a well-known
%multiobjective genetic technique demonstrate that, while not
%outperforming it in terms of pareto dominance, the proposed approach
%can trade the uniformity and granularity qualities of the solutions
%found for obtaining larger Pareto sets, thus representing a further
%choice for designer when objective constrains allow it.

@@

In this work we present PS, a new multiobjective exploration
algorithm. The basic idea is to recognize \emph{interesting} regions, i.e.
regions of the configuration space that are proving to provide better
configurations with regard to other ones. PS evaluates more configurations in
interesting regions while exploring less thoroughly the rest of the
configuration space. After a detailed formal description of the
algorithm and the underlying concepts, we show a 
case study involving the hardware/software exploration of a VLIW
architecture. Qualitative and quantitative comparisons of PS
against a well-known multiobjective genetic approach demonstrate that,
while not outperforming it in terms of pareto dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

@@

\secL{Conclusions}
In this work we presented PS, a multiobjective strategy which 
introduces the concept of Parameter Space representation of Pareto
Front. A case study of a VLIW architecture involving strong
hardware/software dependencies has been analyzed to evaluate
effectiveness with different amount of simulations budget.
A qualitative/quantitative comparison against a widespread multiobjective genetic
approach showed how the proposed strategy can result in a different
distribution of Pareto points which trades solutions granularity with
improved Pareto extension.

@@

>>>>

\newcommand{\ee}{{\sf EPIC-Explorer}}

@@

%\maketitle
\begin{abstract}
%Embedded Systems Design requires conflicting objectives to be
%optimized with an appropriate choice of hardware/software parameters.
%In this work we present PS, a new multi-objective exploration algorithm
%exploiting the parameter space representation of the pareto front.
%After a detailed formal description of the algorithm and the
%underlying concepts, we show a case study involving the
%hardware/software exploration of a VLIW architecture. Qualitative and
%quantitative comparisons of the PS approach against a well-known
%multi-objective genetic technique demonstrate that, while not
%outperforming it in terms of pareto dominance, the proposed approach
%can trade the uniformity and granularity qualities of the solutions
%found for obtaining larger Pareto sets, thus representing a further
%choice for designer when objective constrains allow it.

@@

In this work we present PS, a new multi-objective optimization algorithm and evaluate it in the context of the exploration the embedded system design space. The basic idea is to recognize \emph{interesting} regions, i.e.
regions of the configuration space that are proving to provide better
configurations with regard to other ones. PS evaluates more configurations in
interesting regions while exploring less thoroughly the rest of the
configuration space. After a detailed formal description of the
algorithm and the underlying concepts, we show a 
case study involving the hardware/software exploration of a VLIW
architecture. Qualitative and quantitative comparisons of PS
against a well-known multi-objective genetic approach demonstrate that,
while not outperforming it in terms of pareto dominance, the proposed
approach can trade the uniformity and granularity qualities of the
solutions found for obtaining larger Pareto sets, thus representing a
further choice for designer when objective constrains allow it.
\end{abstract}

@@

\section{Conclusions}
\secL{Conclusions}
In this work we presented PS, a multi-objective strategy which 
introduces the concept of Parameter Space representation of Pareto
Front. A case study of a VLIW architecture involving strong
hardware/software dependencies has been analyzed to evaluate
effectiveness with different amount of simulations budget.
A qualitative/quantitative comparison against a widespread multi-objective genetic
approach showed how the proposed strategy can result in a different
distribution of Pareto points which trades solutions granularity with
improved Pareto extension.

@@

\input{appendix.tex}
